---
title: "Verbal_Autopsy"
author: "Carmen Chan, Neha Shah, Shuhao Zhou, Shannon Walsh"
date: "3/23/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r}
library("dplyr")
```

```{r}
df_va_original <- read.csv("./IHME_PHMRC_VA_DATA_CHILD_Y2013M09D11_0.csv")
```

**Some of the columns were dropped as they are irrelevant/duplicated or not informative (for example, the frequency of the words doesn't provide much information without a context)**

```{r}
# drop columns that are not necessary
df_va <- subset(df_va_original, select = -c(gs_code34,va34,gs_code46,gs_text46,va46,gs_code55,gs_text55,va55,gs_comorbid1,gs_comorbid2,gs_level,g1_01d,g1_01m,g1_01y,g1_05,g1_06d,g1_06m,g1_06y,g1_07a,g1_07b,g1_07c,g1_08,g1_09,g1_10,g2_01,g2_02,g2_03ad,g2_03am,g2_03ay,g2_03bd,g2_03bm,g2_03by,g2_03cd,g2_03cm,g2_03cy,g2_03dd,g2_03dm,g2_03dy,g2_03ed,g2_03em,g2_03ey,g2_03fd,g2_03fm,g2_03fy,g3_01,g4_02,g4_03a,g4_03b,g4_04,g4_05,g5_05,g5_06a,g5_06b,g5_07,g5_08,word_diseas,word_final,word_child,word_condit,word_digest,word_glucos,word_bodi,word_tetanus,word_hand,word_failur,word_reduc,word_son,word_breath,word_look,word_till,word_spot,word_proper,word_medic,word_found,word_girl,word_ray,word_babi,word_privat,word_reason,word_poison,word_bring,word_renal,word_rash,word_healthi,word_interview,word_acquir,word_accid,word_test,word_cough,word_respiratori,word_mother,word_traffic,word_hospit,word_come,word_abdomen,word_abl,word_result,word_pregnanc,word_suffer,word_check,word_pass,word_famili,word_die,word_road,word_colleg,word_dengu,word_doctor,word_drown,word_tumor,word_cardio,word_eat,word_fall,word_examin,word_acidosi,word_fire,word_fit,word_sepsi,word_nilouf,word_happen,word_head,word_headach,word_skin,word_blood,word_home,word_hypertens,word_immedi,word_bad,word_inject,word_left,word_leg,word_daughter,word_malnutrit,word_milk,word_clock,word_anemia,word_nilof,word_complain,word_hour,word_nurs,word_snake,word_pain,word_pneumonia,word_polic,word_provid,word_boy,word_recov,word_malaria,word_asthma,word_explain,word_scan,word_gandhi,word_born,word_lung,word_stomach,word_difficulti,word_weak,word_client,word_time,word_told,word_transfus,word_treat,word_unconsci,word_water,word_loos,word_week,word_stool,word_ill,word_lot,word_jaundic,word_communiti,word_health,word_deliv,word_drink,word_servic,word_fine,word_eye,word_particip,word_money,word_chest,word_increas,word_live,word_expir,word_normal,word_brain,word_stay,word_urin,word_remov,word_admit,word_bite,word_center,word_measl,word_kept,word_especi,word_neck,word_serious,word_due,word_care,word_day,word_pox,word_hiv,word_icu,word_start,word_nose,word_leukemia,word_caus,word_near,word_morn,word_vomit,word_accord,word_gastric,word_receiv,word_coma,word_father,word_clinic,word_emerg,word_month,word_birth,word_treatment,word_sick,word_dehydr,word_prescrib,word_children,word_motion,word_refer,word_ward,word_certif,word_advis,word_hous,word_medicin,word_play,word_heart,word_diarrhea,word_baby,word_mouth,word_sever,word_shock,word_dead,word_oper,word_night,word_indraw,word_provinci,word_cancer,word_brought,word_even,word_convuls,word_addit,word_deceas,word_take,word_oxygen,word_infect,word_cold,word_misplac,word_swell,word_respond,word_transfer,word_thank,word_cri,word_sudden,word_continu,word_sent,word_stop,word_get,word_fever,word_notic,word_hole,word_kidney,word_bluish,word_yellow,word_injuri,word_pulmonari,c6_11,c6_12,c6_13,c6_14,newid,g5_02,module))
```

```{r}
str(df_va)
```

**transform data type: change the characters into categorical variables**

```{r}
df_va[sapply(df_va, is.character)] <- lapply(df_va[sapply(df_va, is.character)], 
                                                           as.factor)
```

**Compute table distribution for all the categorical variables**

```{r}
for (i in 1:ncol(df_va)){
  if (is.factor(df_va[,i])){
    table <- data.frame(table(df_va[,i]))
    colnames(table) <- c(names(df_va)[i],"Freq")
    print(table)
    }
}
```

**Columns further need to be removed**\
module: remove (same value for all observations)\
c1_02: remove, as only applies to multiple births (info is covered by c1_01 already)\
c1_04: remove, as it only applies to cases where moms were dead (and info is covered in c1_03) c1_05: remove, as only applies to cases where moms were dead c1_08a: remove since only showing units (may refer to c1_08b for detailed values)\
c1_10,c1_10d,c1_10m,c1_10y: remove, duplicated with g5_01 c1_11: remove, since we only want to focus on children who shouldn't have been dead at birth\
c1_15: remove, since there's only one class\
c1_16: remove, since there's only one class\
c1_17: remove, since there's only one class\
c1_18: remove, since there's only one class\
c1_19_1 to c1_19_6: remove, since there's only one class\
c1_24: remove, since only showing units\
c1_24d,c1_24m,c1_24y: remove, duplicated with g5_03\
c1_26: remove, since there's only one class\
c4_07a: remove, since only showing units\
c4_31_1: remove, since only applies to those with rash (information already there in c4_30)\
c4_31_2: remove, as only applies to the ones who developed rash (info already covered in c4_30)\
c4_32: remove, as only applies to the ones who developed rash (info already covered in c4_30)\
c4_45: remove, since most of the value missing\
c4_47_8b: remove, since most of the value missing\
c5_02_13: remove, since there's only one class\
c5_02_14: remove, since there's only one class c5_06_2m,c5_06_2d,c5_06_2y: remove, most of the data is don't know/missing c5_08m,c5_08d,c5_08y: remove, most of the data is don't know/missing

```{r}
df_va <- subset(df_va,select=-c(c1_02,c1_04,c1_05,c1_08a,c1_10,c1_10d,c1_10m,c1_10y,c1_11,c1_15,c1_16,c1_17,c1_18,c1_19_1,c1_19_2,c1_19_3,c1_19_4a,c1_19_4b,c1_19_5,c1_19_6,c1_24,c1_24d,c1_24m,c1_24y,c1_26,c4_07a,c4_31_1,c4_31_2,c4_32,c4_45,c4_47_8b,c5_02_13,c5_02_14,c5_06_2m,c5_06_2d,c5_06_2y,c5_08m,c5_08d,c5_08y))
```

**Columns need further processed**\
g5_01d,g5_01m,g5_01y: compute DOB and then remove\
g5_03d,g5_03m,g5_03y: compute DOD and then remove

```{r}
# Compute DOB
df_va$DOB_month <- match(df_va$g5_01m,month.name)
df_va$DOB_str <- paste(df_va$g5_01y,"-",df_va$DOB_month,"-",df_va$g5_01d)
df_va$DOB <- as.POSIXct(df_va$DOB_str, format="%Y - %m - %d",tz="UTC")
df_va <- subset(df_va,select=-c(DOB_str,DOB_month))
```

```{r}
# Compute DOD
df_va$DOD_month <- match(df_va$g5_03m,month.name)
df_va$DOD_str <- paste(df_va$g5_03y,"-",df_va$DOD_month,"-",df_va$g5_03d)
df_va$DOD <- as.POSIXct(df_va$DOD_str, format="%Y - %m - %d",tz="UTC")
df_va <- subset(df_va,select=-c(DOD_str,DOD_month))
```

```{r}
df_va <- subset(df_va,select=-c(g5_01d,g5_01m,g5_01y,g5_03d,g5_03m,g5_03y))
```

c1_22a: combine "hospital" with "other health facility" to "Health facility"

```{r}
df_va$c1_22a <- as.character(df_va$c1_22a)
df_va$c1_22a[df_va$c1_22a=="Hospital"] <- "Health Facility"
df_va$c1_22a[df_va$c1_22a=="Other Health Facility"] <- "Health Facility"
df_va$c1_22a <- as.factor(df_va$c1_22a)
```

c5_06_1d,c5_06_1m,c5_06_1y: remove, as most of the records are missing

```{r}
df_va <- subset(df_va,select=-c(c5_06_1d,c5_06_1m,c5_06_1y))
```

**Imputation and removal for missing values**\
Check missing values for numeric variables

```{r}
df_va_num <- select_if(df_va,is.numeric)
```

```{r}
summary(df_va_num)
```

Remove c5_07_1 and c5_07_1 since most of them are 0 (missing in this case)

```{r}
df_va <- subset(df_va,select=-c(c5_07_1,c5_07_2))
```

c1_08b: re-code "9999" as NA

```{r}
df_va$c1_08b[df_va$c1_08b==9999] <- NA
```

Drop g5_04a,g5_04b,g5_04c and use DOB and DOD to compute the age of death (in years)

```{r}
df_va <- subset(df_va,select=-c(g5_04a,g5_04b,g5_04c))
df_va$age_death <- as.numeric(difftime(df_va$DOD, df_va$DOB, units = "days"))/365
```

```{r}
library(naniar)
vis_miss(df_va_num)
```

split training and test set

```{r}
library(caTools)
set.seed(123)
split = sample.split(df_va$site, SplitRatio = 0.7)
training_set = subset(df_va, split == TRUE)
test_set = subset(df_va, split == FALSE)
```

Deal with "Don't know", "Refuse to answer" and original missing value for all the categorical variables\
1) If the sum of counts of all of them three is \>=10: re-code all of them with "Don't know" as a group itself\
2) If the sum of counts of all of them three is \<10: re-code all of them as missing value

```{r}
# Define function impute_cat
impute_cat <- function(df_va){
  cat_num=c()
  for (i in 1:ncol(df_va)){
    if(is.factor(df_va[,i])){
      cat_num <- append(cat_num,i)} # Compute the positions of the columns where the variable is categorical
    }
  for (i in cat_num){
    count_dk <- sum(df_va[,i]=="Don't Know",na.rm=TRUE)
    count_rta <- sum(df_va[,i]=="Refused to Answer",na.rm=TRUE)
    count_missing <- sum(is.na(df_va[,i]),na.rm=FALSE)
    count_whitespace <- sum(df_va[,i]=="",na.rm=TRUE)
    count_all <- count_dk+count_rta+count_missing+count_whitespace
    df_va[,i] <- as.character(df_va[,i])
    if (count_all==10||count_all>15){
      if (count_rta>0){
        df_va[,i][df_va[,i]=="Refused to Answer"] <- "Don't Know"
        }
      if (count_whitespace>0){
        df_va[,i][df_va[,i]==""] <- "Don't Know"
        }
      if (count_missing>0){
        df_va[,i][is.na(df_va[,i])] <- "Don't Know"
        }
      }else{
        if (count_rta>0){
          df_va[,i][df_va[,i]=="Refused to Answer"] <- NA
          }
        if (count_dk>0){
          df_va[,i][df_va[,i]=="Don't Know"] <- NA
          }
        if (count_whitespace>0){
          df_va[,i][df_va[,i]==""] <- NA
        }
      }
    df_va[,i] <- as.factor(df_va[,i])
  }
  return(df_va)
}

```

```{r}
# apply impute_cat on both training set and test set
training_set <- impute_cat(training_set)
test_set <- impute_cat(test_set)
```

Impute missing values of c1_08b with mean of the data by different site groups (fit on training set and transform both sets)

```{r}
tapply(training_set$c1_08b,training_set$site, mean, na.rm=TRUE)
```

```{r}
correct_weight <- function(x1, x2){
  if(is.na(x1)){
    if(x2=="AP"){
      return(2710)
      }
    else if(x2=="Bohol"){
      return(2972)
      }
    else if (x2=="Dar"){
      return(2964)
    }
    else if (x2=="Mexico"){
      return(2555)
    }
    else if (x2=="Pemba"){
      return(3058)
    }
    else{return(2716)}
  }else{
    return(x1)}
}
```

```{r}
training_set$c1_08b <- apply(training_set[,c("c1_08b","site")], 1, function(x) correct_weight(x[1],x[2]))
test_set$c1_08b <- apply(test_set[,c("c1_08b","site")], 1, function(x) correct_weight(x[1],x[2]))
```

**Remove all the remaining missing data for both sets**

```{r}
training_set <- training_set[complete.cases(training_set),]
test_set <- test_set[complete.cases(test_set),]
```

=======


**Visualization for EDA**
```{r}
# Change colname names for training_set_copy
library(ggpubr)
library(plyr)
training_set_copy <- training_set
```


```{r}
colnames(training_set_copy)<- mapvalues(colnames(training_set_copy), from=c("g4_06","g4_07","g4_08"), to=c("num_people_live_at_address","num_rooms_in_household","separate_room_for_cooking"))
colnames(training_set_copy)<- mapvalues(colnames(training_set_copy), from=c("c1_09","c1_01","c1_03"), to=c("Gender","Singleton_or_Multiple_Birth","Mother_Living_or_Deceased"))
colnames(training_set_copy)<- mapvalues(colnames(training_set_copy), from=c("c1_06a","c1_07","c1_08b","c1_12","c1_13"), to=c("Location_of_Birth","Size_at_Birth","Weight_at_Birth","Did_the_Baby_Cry","Did_the_Baby_Move"))
colnames(training_set_copy)<- mapvalues(colnames(training_set_copy), from=c("num_people_live_at_address","num_rooms_in_household","separate_room_for_cooking"), to=c("Num_People_Live_at_Address","Num_Rooms_in_Household","Separate_Room_for_Cooking"))
colnames(training_set_copy)<- mapvalues(colnames(training_set_copy), from=c("c1_14","c1_20","c1_21","c1_22a","c1_25","c4_01","c4_02","c4_03"), to=c("Did_the_Baby_Breathe","Age_at_Onset_of_Illness","Duration_of_Illness", "Location_of_Death","Age_at_Time_of_Death","Fever_During_Illness","Duration_of_Fever_in_Days","Did_the_Fever_Continue_to_Death"))
colnames(training_set_copy)<- mapvalues(colnames(training_set_copy), from=c("c4_04","c4_05","c4_06","c4_07b","c4_08","c4_09","c4_10","c4_11","c4_12","c4_13","c4_14","c4_15","c4_16","c4_17","c4_18", "c4_19"), to=c("Severity_of_Fever","Fever_Pattern","Loose_Liquid_Stool", "Highest_Num_Loose_Stool_per_Day_During_Illness","Num_Days_Before_Death_Loose_Stool_Began","Loose_Stool_Cont_Until_Death","Num_Days_Before_Death_Loose_Stool_Stopped","Blood_in_Stool", "Cough_During_Illness", "Duration_of_Cough", "Severity_of_Cough", "Vomitus_after_Coughing", "Difficulty_Breathing", "Duration_of_Difficulty_Breathing", "Fast_Breathing", "Duration_of_Fast_Breathing"))
colnames(training_set_copy)<- mapvalues(colnames(training_set_copy), from=c("c4_20", "c4_22", "c4_23", "c4_24", "c4_25", "c4_26","c4_27", "c4_28", "c4_29", "c4_30", "c4_33", "c4_34", "c4_35", "c4_36", "c4_37", "c4_38", "c4_39","c4_40", "c4_41", "c4_42"), to=c("Indrawing_of_Chest", "Breathing_Stridor", "Breathing_Grunting", "Breathing_Wheezing", "Convulsions", "Loss_of_Consciousness", "Duration_Before_Death_LOC_Occurred", "Stiff_Neck", "Bulging_Fontanelle", "Skin_Rash", "Duration_of_Rash", "Blisters_Present_in_Rash", "Limbs_Become_Thin", "Swollen_Legs_or_Feet", "Duration_of_Swelling", "Skin_Flake_Off_in_Patches", "Hair_Color_Change_to_Red_Yellow", "Protruding_Belly", "Pallor_or_Lack_of_Blood", "Swelling_in_Armpits"))
colnames(training_set_copy)<- mapvalues(colnames(training_set_copy), from=c("c4_43","c4_44","c4_46", "c4_47_1","c4_47_2", "c4_47_3", "c4_47_4", "c4_47_5","c4_47_6","c4_47_7","c4_47_8a","c4_47_9","c4_47_10","c4_47_11"), to=c("Whitish_Rash_in_Mouth", "Bleeding_Seen","Skin_Turned_Black", "Suffered_Road_Traffic_Injury", "Suffered_a_Fall", "Suffered_Drowning", "Suffered_Poisoning", "Suffered_Bite_Sting", "Suffered_Burn_Fire", "Victim_of_Violence", "Other_Injury", "Unsure_if_Injury_Occurred","Refused_to_Answer_if_Deceased_Suffered_Injury", "Did_Not_Suffer_Injury" ))
colnames(training_set_copy)<- mapvalues(colnames(training_set_copy), from=c("c4_48","c4_49","c5_01","c5_02_1","c5_02_2","c5_02_3","c5_02_4","c5_02_5","c5_02_6","c5_02_7","c5_02_8","c5_02_9", "c5_02_10", "c5_02_11a","c5_02_12","c5_04", "c5_05","c5_10", "c5_11", "c5_17", "c5_18", "c5_19"), to=c("Injury_Intentionally_Inflicted_by_Someone","Days_Survived_After_Injury","Sought_Care_While_Ill", "Care_Sought_Traditional_Healer", "Care_Sought_Homeopath", "Care_Sought_Religious_Leader", "Care_Sought_Governmental_Hospital", "Care_Sought_Governmental_Health_Center_Clinic", "Care_Sought_Private_Hospital", "Care_Sought_Community_Based_Practioner", "Care_Sought_Trained_Birth_Attendant", "Care_Sought_Private_Physician", "Care_Sought_Pharmacy", "Care_Sought_Other_Provider", "Care_Sought_Relative_Friend", "Health_Records_for_Deceased", "Granted_Access_to_Health_Records", "Death_Certificate_Issued", "Granted_Access_to_Death_Certificate", "Mother_Ever_Tested_for_HIV", "Mother_HIV_Positive", "Mother_AIDS_Positive"))
```


Rename test set
```{r}
colnames(test_set)<- mapvalues(colnames(test_set), from=c("g4_06","g4_07","g4_08"), to=c("num_people_live_at_address","num_rooms_in_household","separate_room_for_cooking"))
colnames(test_set)<- mapvalues(colnames(test_set), from=c("c1_09","c1_01","c1_03"), to=c("Gender","Singleton_or_Multiple_Birth","Mother_Living_or_Deceased"))
colnames(test_set)<- mapvalues(colnames(test_set), from=c("c1_06a","c1_07","c1_08b","c1_12","c1_13"), to=c("Location_of_Birth","Size_at_Birth","Weight_at_Birth","Did_the_Baby_Cry","Did_the_Baby_Move"))
colnames(test_set)<- mapvalues(colnames(test_set), from=c("num_people_live_at_address","num_rooms_in_household","separate_room_for_cooking"), to=c("Num_People_Live_at_Address","Num_Rooms_in_Household","Separate_Room_for_Cooking"))
colnames(test_set)<- mapvalues(colnames(test_set), from=c("c1_14","c1_20","c1_21","c1_22a","c1_25","c4_01","c4_02","c4_03"), to=c("Did_the_Baby_Breathe","Age_at_Onset_of_Illness","Duration_of_Illness", "Location_of_Death","Age_at_Time_of_Death","Fever_During_Illness","Duration_of_Fever_in_Days","Did_the_Fever_Continue_to_Death"))
colnames(test_set)<- mapvalues(colnames(test_set), from=c("c4_04","c4_05","c4_06","c4_07b","c4_08","c4_09","c4_10","c4_11","c4_12","c4_13","c4_14","c4_15","c4_16","c4_17","c4_18", "c4_19"), to=c("Severity_of_Fever","Fever_Pattern","Loose_Liquid_Stool", "Highest_Num_Loose_Stool_per_Day_During_Illness","Num_Days_Before_Death_Loose_Stool_Began","Loose_Stool_Cont_Until_Death","Num_Days_Before_Death_Loose_Stool_Stopped","Blood_in_Stool", "Cough_During_Illness", "Duration_of_Cough", "Severity_of_Cough", "Vomitus_after_Coughing", "Difficulty_Breathing", "Duration_of_Difficulty_Breathing", "Fast_Breathing", "Duration_of_Fast_Breathing"))
colnames(test_set)<- mapvalues(colnames(test_set), from=c("c4_20", "c4_22", "c4_23", "c4_24", "c4_25", "c4_26","c4_27", "c4_28", "c4_29", "c4_30", "c4_33", "c4_34", "c4_35", "c4_36", "c4_37", "c4_38", "c4_39","c4_40", "c4_41", "c4_42"), to=c("Indrawing_of_Chest", "Breathing_Stridor", "Breathing_Grunting", "Breathing_Wheezing", "Convulsions", "Loss_of_Consciousness", "Duration_Before_Death_LOC_Occurred", "Stiff_Neck", "Bulging_Fontanelle", "Skin_Rash", "Duration_of_Rash", "Blisters_Present_in_Rash", "Limbs_Become_Thin", "Swollen_Legs_or_Feet", "Duration_of_Swelling", "Skin_Flake_Off_in_Patches", "Hair_Color_Change_to_Red_Yellow", "Protruding_Belly", "Pallor_or_Lack_of_Blood", "Swelling_in_Armpits"))
colnames(test_set)<- mapvalues(colnames(test_set), from=c("c4_43","c4_44","c4_46", "c4_47_1","c4_47_2", "c4_47_3", "c4_47_4", "c4_47_5","c4_47_6","c4_47_7","c4_47_8a","c4_47_9","c4_47_10","c4_47_11"), to=c("Whitish_Rash_in_Mouth", "Bleeding_Seen","Skin_Turned_Black", "Suffered_Road_Traffic_Injury", "Suffered_a_Fall", "Suffered_Drowning", "Suffered_Poisoning", "Suffered_Bite_Sting", "Suffered_Burn_Fire", "Victim_of_Violence", "Other_Injury", "Unsure_if_Injury_Occurred","Refused_to_Answer_if_Deceased_Suffered_Injury", "Did_Not_Suffer_Injury" ))
colnames(test_set)<- mapvalues(colnames(test_set), from=c("c4_48","c4_49","c5_01","c5_02_1","c5_02_2","c5_02_3","c5_02_4","c5_02_5","c5_02_6","c5_02_7","c5_02_8","c5_02_9", "c5_02_10", "c5_02_11a","c5_02_12","c5_04", "c5_05","c5_10", "c5_11", "c5_17", "c5_18", "c5_19"), to=c("Injury_Intentionally_Inflicted_by_Someone","Days_Survived_After_Injury","Sought_Care_While_Ill", "Care_Sought_Traditional_Healer", "Care_Sought_Homeopath", "Care_Sought_Religious_Leader", "Care_Sought_Governmental_Hospital", "Care_Sought_Governmental_Health_Center_Clinic", "Care_Sought_Private_Hospital", "Care_Sought_Community_Based_Practioner", "Care_Sought_Trained_Birth_Attendant", "Care_Sought_Private_Physician", "Care_Sought_Pharmacy", "Care_Sought_Other_Provider", "Care_Sought_Relative_Friend", "Health_Records_for_Deceased", "Granted_Access_to_Health_Records", "Death_Certificate_Issued", "Granted_Access_to_Death_Certificate", "Mother_Ever_Tested_for_HIV", "Mother_HIV_Positive", "Mother_AIDS_Positive"))
```


merge training_set_copy and test_set just for EDA visualization
```{r}
df_merge <- rbind(training_set_copy,test_set)
```



```{r}
library(ggplot2)
```


Plot distribution of all the numerical variables with histogram
```{r}
hist_list <- list()
z <- 0
for (i in 1:ncol(df_merge)){
  if(is.numeric(df_merge[,i])){
    z <- z+1
    hist <- ggplot(data=df_merge,aes_string(names(df_merge)[i]))+
      geom_histogram(fill='lightblue',color='black')+
      labs(x=colnames(df_merge)[i],y= "Frequency")+
      theme(axis.text=element_text(size=12),
        axis.title=element_text(size=8))
    hist_list[[z]] <- hist
  }
}
```


```{r}
ggarrange(plotlist=hist_list,ncol = 2)
```


Plot bar charts for all the categorical variables
```{r}
library(scales)
bar_list <- list()
z <- 0
for (i in 1:ncol(df_merge)){
  if(is.factor(df_merge[,i])){
    z <- z+1
    class <- df_merge %>% 
     group_by((df_merge)[i]) %>%
     dplyr::summarise(number=n()) %>%
     mutate(percentage=percent(number/sum(number))) %>% 
     mutate(pos = cumsum(number)- number/1.5) #Create a variable to indicate the position of labels 
    bar <- ggplot(data=class,aes_string(x=colnames(class)[1],y=colnames(class)[2]))+
      geom_bar(fill='lightblue',color='black',stat = "identity")+
      labs(x=colnames(df_merge)[i],y= "Count")+
      geom_text(aes(label=paste(number,"(",percentage,")")),vjust=-0.5,size=2,color="red")+
      theme(axis.text=element_text(size=8,angle=45,hjust=1),axis.title=element_text(size=8))
    bar_list[[z]] <- bar
  }
}
```


```{r}
# Cause of death have many bins so I plot it separately
bar_list[[2]]
ggarrange(plotlist=bar_list[-2],ncol=2)
```


```{r}
# Top 5 cause of death for the whole dataset: Pneumonia, Diarrhea/Dysentery,Other Defined Causes of Child Deaths, Sepsis, Malaria
top_5 <- c("Pneumonia","Diarrhea/Dysentery","Other Defined Causes of Child Deaths", "Sepsis", "Malaria")
df_merge_top5 <- df_merge[df_merge$gs_text34 %in% top_5,]
```

```{r}
# Remove classes with 0 cases
for (i in 1:length(names(df_merge_top5))){
  if (is.factor(df_merge_top5[,i])){
    df_merge_top5[,i] <- as.character(df_merge_top5[,i])
    df_merge_top5[,i] <- as.factor(df_merge_top5[,i])
  }
}
```

```{r}
table(df_merge_top5$gs_text34)
```


We may also want to see the top 5 cause of death of each area
```{r}
top_5_by_site <- df_merge[,c("site","gs_text34")]%>%
  group_by(site,gs_text34)%>%
  dplyr::summarise(count_cause=n())%>%
  arrange(site,desc(count_cause))%>%
  filter(row_number()==1:5)
  
```

```{r}
top_5_by_site 
```

```{r}
var_list <- names(df_merge_top5)
var_list
```



Plot grouped box plots for numerical variables with top 5 cause of death (across the whole dataset)
```{r fig.width = 10}
var_list <- names(df_merge_top5)
grouped_box_list <- list()
z <- 0
for (i in 3:length(var_list)){
  if (is.numeric(df_merge_top5[,i])){
    z <- z+1
    grouped_box <- ggplot(data=df_merge_top5,aes_string(x=var_list[1],y=var_list[i],fill=var_list[2]))+
      geom_boxplot()+
      theme(axis.text=element_text(size=15,angle=45,hjust=1),axis.title=element_text(size=15),legend.key.size=unit(2,"cm"),legend.title=element_text(size=15),legend.text=element_text(size=15))
    grouped_box_list[[z]] <- grouped_box
  }
}
```

```{r fig.width = 10}
grouped_box_list
```

Plot grouped bar charts for categorical variables with the top 5 cause of death (across the whole dataset) (I only plot for "Pneumonia","Diarrhea/Dysentery","Sepsis")
```{r}
library(dplyr)
top_5 <-c("Pneumonia","Diarrhea/Dysentery","Other Defined Causes of Child Deaths","Sepsis", "Malaria")
```


Compute characteristic/symptom prevalence by site. The number will be used to define if a given characteristic/symptom is more prevalent in one CoD group than another across different sites (example using "Gender" is given below).
```{r}
cause_prev_by_site <- df_merge[,c("site","Gender")]%>%
  group_by(site,Gender)%>%
  dplyr::summarise(count_cause=n())%>%
  dplyr::mutate(prevalence=count_cause/sum(count_cause))%>%
  arrange(site,desc(count_cause))
cause_prev_by_site
```



```{r fig.width = 7}
# Pneumonia
grouped_bar_Pneumonia <- list()
z <- 0
for (i in 3:length(names(df_merge))){
  if (is.factor(df_merge[,i]) & names(df_merge)[i]!="site" & names(df_merge)[i]!="gs_text34"){
    z <- z+1
    grouped_class <- df_merge[df_merge$gs_text34=="Pneumonia",c(1,i)]%>%
      group_by_all()%>%
     dplyr::summarise(count=n())%>%
      group_by(site)%>%
      dplyr::mutate(perc=count/sum(count))
    grouped_bar <- ggplot(data=grouped_class, aes_string(x=names(grouped_class)[2], y=names(grouped_class)[4],fill=names(grouped_class)[2]))+geom_bar(position = 'dodge', stat ='identity')+labs(x=names(grouped_class)[2], y ="perc")+ ggtitle(labs(title = "Cause of death: Pneumonia"))+ theme(text = element_text(size=12),axis.title=element_text(size=12),axis.text=element_text(size=12,hjust=1,angle = 45))+ geom_text(aes(label = scales::percent(perc),y = perc,size=10),position=position_dodge(width=1), vjust =-0.5,size=4)+ facet_grid(. ~ site,scales='free')
    grouped_bar_Pneumonia[[z]] <- grouped_bar
  }
}
grouped_bar_Pneumonia
```

```{r fig.width = 7}
# "Diarrhea/Dysentery"
grouped_bar_Diarrhea <- list()
z <- 0
for (i in 3:length(names(df_merge))){
  if (is.factor(df_merge[,i]) & names(df_merge)[i]!="site" & names(df_merge)[i]!="gs_text34"){
    z <- z+1
    grouped_class <- df_merge[df_merge$gs_text34=="Diarrhea/Dysentery",c(1,i)]%>%
      group_by_all()%>%
     dplyr::summarise(count=n())%>%
      group_by(site)%>%
      dplyr::mutate(perc=count/sum(count))
    grouped_bar <- ggplot(data=grouped_class, aes_string(x=names(grouped_class)[2], y=names(grouped_class)[4],fill=names(grouped_class)[2]))+geom_bar(position = 'dodge', stat ='identity')+labs(x=names(grouped_class)[2], y ="perc")+ ggtitle(labs(title = "Cause of death: Diarrhea/Dysentery"))+ theme(text = element_text(size=12),axis.title=element_text(size=12),axis.text=element_text(size=12,hjust=1,angle = 45))+ geom_text(aes(label = scales::percent(perc),y = perc,size=10),position=position_dodge(width=1), vjust =-0.5,size=4)+ facet_grid(. ~ site,scales='free')
    grouped_bar_Pneumonia[[z]] <- grouped_bar
  }
}
grouped_bar_Pneumonia
```


```{r fig.width = 7}
# Sepsis
grouped_bar_sepsis <- list()
z <- 0
for (i in 3:length(names(df_merge))){
  if (is.factor(df_merge[,i]) & names(df_merge)[i]!="site" & names(df_merge)[i]!="gs_text34"){
    z <- z+1
    grouped_class <- df_merge[df_merge$gs_text34=="Sepsis",c(1,i)]%>%
      group_by_all()%>%
     dplyr::summarise(count=n())%>%
      group_by(site)%>%
      dplyr::mutate(perc=count/sum(count))
    grouped_bar <- ggplot(data=grouped_class, aes_string(x=names(grouped_class)[2], y=names(grouped_class)[4],fill=names(grouped_class)[2]))+geom_bar(position = 'dodge', stat ='identity')+labs(x=names(grouped_class)[2], y ="perc")+ ggtitle(labs(title = "Cause of death: Sepsis"))+ theme(text = element_text(size=12),axis.title=element_text(size=12),axis.text=element_text(size=12,hjust=1,angle = 45))+ geom_text(aes(label = scales::percent(perc),y = perc,size=10),position=position_dodge(width=1), vjust =-0.5,size=4)+ facet_grid(. ~ site,scales='free')
    grouped_bar_sepsis[[z]] <- grouped_bar
  }
}
grouped_bar_sepsis
```






**EDA Visualization of Gender of Deceased**
```{r}
gen <- ggplot(data = training_set_copy) + 
  geom_bar(aes(x = Gender, fill = Gender), stat = "Count")
print(gen)
```

